\documentclass[USenglish]{ifimaster}

\usepackage[export]{adjustbox}
\usepackage[utf8]{inputenc}
\usepackage{ifimasterforside}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{floatrow}
\usepackage{rotating}
\usepackage{pdflscape}

\usepackage[acronym, toc]{glossaries}
\usepackage[toc,page]{appendix}

\usepackage[backend=biber,sorting=none]{biblatex}
\usepackage{listings}

\usepackage{tabularx}
\usepackage{cleveref}

% For listings
\usepackage{bera}
\usepackage{listings}
\usepackage{xcolor}
\colorlet{punct}{red!60!black}
\definecolor{background}{HTML}{EEEEEE}
\definecolor{delim}{RGB}{20,105,176}
\colorlet{numb}{magenta!60!black}
\lstdefinelanguage{json}{
    basicstyle=\normalfont\ttfamily,
    numbers=left,
    numberstyle=\scriptsize,
    stepnumber=1,
    numbersep=8pt,
    showstringspaces=false,
    breaklines=true,
    frame=lines,
    backgroundcolor=\color{background},
    literate=
     *{0}{{{\color{numb}0}}}{1}
      {1}{{{\color{numb}1}}}{1}
      {2}{{{\color{numb}2}}}{1}
      {3}{{{\color{numb}3}}}{1}
      {4}{{{\color{numb}4}}}{1}
      {5}{{{\color{numb}5}}}{1}
      {6}{{{\color{numb}6}}}{1}
      {7}{{{\color{numb}7}}}{1}
      {8}{{{\color{numb}8}}}{1}
      {9}{{{\color{numb}9}}}{1}
      {:}{{{\color{punct}{:}}}}{1}
      {,}{{{\color{punct}{,}}}}{1}
      {\{}{{{\color{delim}{\{}}}}{1}
      {\}}{{{\color{delim}{\}}}}}{1}
      {[}{{{\color{delim}{[}}}}{1}
      {]}{{{\color{delim}{]}}}}{1},
}

\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}
\definecolor{pred}{rgb}{0.9,0,0}
\definecolor{pgrey}{rgb}{0.46,0.45,0.48}
\lstset{language=Java,
  showspaces=false,
  showtabs=false,
  breaklines=true,
  showstringspaces=false,
  breakatwhitespace=true,
  commentstyle=\color{pgreen},
  keywordstyle=\color{pblue},
  stringstyle=\color{pred},
  basicstyle=\ttfamily,
  moredelim=[il][\textcolor{pgrey}]{$$},
  moredelim=[is][\textcolor{pgrey}]{\%\%}{\%\%}
}


\makeglossaries

% Acronyms
\newacronym{acm}{ACM}{Agile Computing Middleware}
\newacronym{afro}{AFRO}{Adaption Framework foR Web Services prOvision}
\newacronym{amqp}{AMQP}{Advanced Message Queuing Protocol}
\newacronym{api}{API}{Application Program Interface}
\newacronym{coap}{CoAP}{The Constrained Application Protocol}
\newacronym{cnr}{CNR}{Combat Net Radio}
\newacronym{cots}{COTS}{Commercial off-the-shelf}
\newacronym{dsproxy}{DSProxy}{Delay and disruption tolerant SOAP Proxy}
\newacronym{dil}{DIL}{Disconnected, Intermittent and Limited}
\newacronym{edge}{EDGE}{Enhanced Data rates for GSM Evolution}
\newacronym{efx}{EFX}{Efficient XML}
\newacronym{ffi}{FFI}{Norwegian Defence Research Establishment}
\newacronym{fec}{FEC}{Forward Error Correction}
\newacronym{ftp}{FTP}{File Transfer Protocol}
\newacronym{http}{HTTP}{Hypertext Transfer Protocol}
\newacronym{ietf}{IETF}{Internet Engineering Task Force}
\newacronym{iot}{IoT}{Internet of Things}
\newacronym{ip}{IP}{Internet Protocol}
\newacronym{json}{JSON}{JavaScript Object Notation}
\newacronym{jvm}{JVM}{Java Virtual Machine}
\newacronym{kda}{KDA}{Kongsberg Defence \& Aerospace}
\newacronym{los}{LOS}{Line of Sight}
\newacronym{lte}{LTE}{Long-Term Evolution}
\newacronym{manet}{MANET}{Mobile ad hoc network}
\newacronym{mockets}{Mockets}{Mobile Sockets}
\newacronym{mtu}{MTU}{Maximum Transfer Unit}
\newacronym{nato}{NATO}{North Atlantic Treaty Organization}
\newacronym{nec}{NEC}{Network Enabled Capability}
\newacronym{netem}{NetEm}{Network Emulator}
\newacronym{nffi}{NFFI}{NATO Friendly Force Information}
\newacronym{nio}{NIO}{Java new/non-blocking I/O}
\newacronym{ntnu}{NTNU}{Norwegian University of Science and Technology}
\newacronym{oasis}{OASIS}{Organization for the Advancement of Structured
Information Standards}
\newacronym{per}{PER}{Packet Error Rate}
\newacronym{qos}{QoS}{Quality of Service}
\newacronym{rest}{REST}{Representational State Transfer}
\newacronym{satcom}{SATCOM}{Satellite Communication}
\newacronym{sctp}{SCTP}{Stream Control Transmission Protocol}
\newacronym{soa}{SOA}{Service Oriented Architecture}
\newacronym{soap}{SOAP}{SOAP}
\newacronym{std}{STD}{Standard Deviation}
\newacronym{tbf}{TBF}{Token Bucket Filter}
\newacronym{tcp}{TCP}{Transmission Control Protocol}
\newacronym{udp}{UDP}{User Datagram Protocol}
\newacronym{uri}{URI}{Uniform Resource Identifier}
\newacronym{rtt}{RTT}{Round-trip time}
\newacronym{w3c}{W3C}{World Wide Web Consortium}
\newacronym{wsdl}{WSDL}{Web Services Description Language}
\newacronym{www}{WWW}{World Wide Web}
\newacronym{xml}{XML}{Extensible Markup Language}

\title{Improving the performance of Web Services in Disconnected, Intermittent
and Limited Environments}
\author{Joakim Johanson Lindquister}
\bibliography{references}

\begin{document}
\ififorside{}

\chapter*{Abstract}

Using \gls{cots} software over networks that are \gls{dil} may not perform
satisfactory, or can even break down entirely. Such networks are characterized
by seeing frequent disruptions for both shorter and longer periodes, as well as
high delays, low data rates and high packet error rates. In this thesis, we
design and implement a HTTP proxy to improve the performance of Web services in
DIL environments. The main idea of our design is to deploy a pair of proxies to
facilitate HTTP communication between entities. As an optimization technique, we
evaluate the usage of alternative transport protocols to carry information
across these type of networks.

The implemented proxy is designed to support \gls{http}, \gls{amqp} and
\gls{coap} for \textit{inter-proxy communication}. By introducing a proxy pair,
we was able to break the end-to-end dependency between two services
communicating over a DIL network, thus achieving higher reliability.
Furthermore, experimental results reveal that when the message size is low,
\gls{coap} has a lower overhead than the other protocols. MER OM RESULTER
SENERE.



\chapter*{Preface}

This master thesis was written at the Department of Informatics at the Faculty
of Mathematics and Natural Sciences, University at the University of Oslo in
2015/2016. It was written in cooperation with \gls{ffi}, which provided the
thesis topic and supervison.


\tableofcontents

\glsresetall

\include{introduction}
\include{background}
\include{related_work}
\include{requirement_analysis}
\include{design_implementation}
\include{evaluation}

\chapter{Conclusion and Future Work}
\label{chapter:conclusion}

\section{Conclusion}
Revisit problem statement. Ta opp igjen premisser og krav.

2. Oppfylt.

3. Interoperable with standarized solutions. By design. Peke på funksjonstesting.

4. Ihvertfall by design. "Secure Web services". Enten kryptering eller security.

Vise hvordan vi har oppfylt requirements.
Bruk by design om mulig.

Gni inn at goals were achieved. Prototype og anbefaling.


\section{Future Work}

\subsection{Andre protokoller}

\subsection{Improvements of Proxy}

Runtime valg av optimalisering.

\subsection{Known bugs}

HTTP Request headers are also included in the HTTP response.

\subsubsection{HTTP}

Not necessary to use proxy message here, since we can simply forward the HTTP request.

IPSEC.

\pagebreak
\printbibliography{}
\printglossaries{}

\include{appendix}

\end{document}
